---
title: "Tidy Data"
output: github_document
---

```{r echo = FALSE, message = FALSE}
library(tidyverse)
library(readxl)
library(haven) 
```

This document will show how to tidy data.
```{r}
pulse_df = 
  read_sas("data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    cols = bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    values_to = "bdi_score",
    names_prefix = "bdi_score_"
  ) %>% 
  mutate(
    visit = replace(visit, visit == 'bl', '00m')
  ) %>% 
  relocate(id, visit)
```

One More Example.
```{r}
litters_df = 
  read_csv('data/FAS_litters.csv', na = c("NA", ".", "")) %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    cols = gd0_weight:gd18_weight,
    names_to = "gd_time", 
    values_to = "weight"
  ) %>% 
  mutate(
    gd_time = case_match(
      gd_time,
      "gd0_weight" ~ 0,
      "gd18_weight" ~ 18
    )
  )
```
case_match function is useful when there are multiple levels in a column in a string format that you want to change into a numerical variable.

## Pivot wider
Useful when post processing analysis of results

```{r}
analysis_df = 
  tibble(
    group = c("treatment", "treatment", "control", "control"), 
    time = c("pre", "post", "pre", "post"),
    mean = c(4, 10, 4.2, 5)
  )
```

Markdown table formatting using knitr package, kable function. More user-friendly, readable format
```{r}
analysis_df %>% 
  pivot_wider(
    names_from = time,
    values_from = mean
  ) %>% 
  knitr::kable()
```
Lord of the Rings Data: goal is to pull all data in three seperate tables in an excel file into a single data frame.
```{r}
fellowship_ring = 
  read_excel("data/LotR_Words.xlsx", range = "B3:D6") %>% 
  mutate(movie = "fellowship_ring")

two_towers = 
  read_excel("data/LotR_Words.xlsx", range = "F3:H6") %>% 
  mutate(movie = "two_towers")

return_king = 
  read_excel("data/LotR_Words.xlsx", range = "J3:L6") %>% 
  mutate(movie = "return_king")

lotr_df = 
  bind_rows(
    fellowship_ring, two_towers, return_king) %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    col = female:male,
    names_to = 'sex',
    values_to = 'words'
  ) %>% 
  relocate(movie) %>% 
  mutate(race = str_to_lower(race))
```

## Join FAS datasets

Import `litters` dataset.

```{r}
litters_df = 
  read_csv("data/FAS_litters.csv", na = c("NA", ".", "")) %>% 
  janitor::clean_names() %>% 
  mutate(
    wt_gain = gd18_weight - gd0_weight
  ) %>% 
  separate(
    group, into= c("dose", "day_of_treatment"), sep = 3
  )
litters_df

```
Import `pups` next

```{r}
pups_df = 
  read_csv("data/FAS_pups.csv", na = c("NA", ".", "")) %>% 
  janitor::clean_names() %>% 
  mutate(
    sex = case_match(
      sex,
      1 ~ "male",
      2~ "female"
    )
  )
```
Joni the datasets!

```{r}
fas_df = 
  left_join(pups_df, litters_df, join_by(litter_number)) %>% 
  relocate(litter_number, dose, day_of_treatment)

```

