---
title: "20240920_data_manipulation"
output: github_document
---


# Import Library
```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)
```

This document will show how to *manipulate* data.
```{r, message = FALSE}
litters_df = 
  read_csv("data/FAS_litters.csv", na = c("NA", "", "."))

litters_df = janitor::clean_names(litters_df)

pups_df = 
  read_csv("data/FAS_pups.csv", na = c("NA", "", "."))

pups_df = janitor::clean_names(pups_df)
```
## `select`

Use `select` to select variables

```{r}
select(litters_df, group, litter_number, gd0_weight)
```

Select everything from group to gd18_weight column

```{r}
select(litters_df, group:gd18_weight)

```

Select everything BUT pups survive column
```{r}
select(litters_df, -pups_survive)
```

Select and combine the parameter functions mentioned above
```{r}
select(litters_df, -(group: gd18_weight))
```

select only columns with certain characters of string or characters using key word

```{r}
select(litters_df, starts_with("gd"))

select(litters_df, contains("pups"))
```

Select and rename certain column name
```{r}
select(litters_df, GROUP = group)
```

Only renaming column name

```{r}
rename(litters_df, GROUP = group)
```

Putting the column in order of preference and using everything function to call all other columns
  The everything function 
```{r}
select(litters_df, litter_number, gd0_weight, everything())

relocate(litters_df, litter_number, gd0_weight)
```
### Learning Assessment
In the pups data, select the columns containing litter number, sex, and PD ears
```{r}
select(pups_df, litter_number, sex, pd_ears)
```

## `filter`
The `filter` filters out rows. Unlike select, `filter` applies on rows. It can have conditional way to filter out variables of interest

```{r}
filter(litters_df, gd_of_birth == 20)
filter(litters_df, gd_of_birth == 20)
```
```{r}
filter(litters_df, pups_born_alive > 8)
filter(litters_df, pups_born_alive >= 8)
```

Filtering out when values are NOT equal to a certain value
```{r}
filter(litters_df, pups_born_alive != 9)
```

Filtering out when values are equal to certain string
double equals doesn't work when you give list of strings
```{r}
filter(litters_df, group == 'Low8')
filter(litters_df, group %in% c("Low7", "Low8"))
```

Filtering using multiple conditional statments. Both comma and | should be able to separate the conditions.
```{r}
filter(litters_df, group %in% c("Low7", "Low8"), pups_born_alive == 8)

filter(litters_df, group %in% c("Low7", "Low8") | pups_born_alive == 8)
```

DROP NA: filtering out rows with missing variables
```{r}
drop_na(litters_df)

drop_na(litters_df, gd0_weight)

```
If you want to overwrite dropped NA values variables, must save change by redefining variables.

### Learning Assessment
```{r}
filter(pups_df, sex == 1)

filter(pups_df, pd_walk<11, sex<2)
```

## `mutate`

Apply arithmetic using certain columns and can create new columns to add into the data frame or overwrite variables in certain column to apply string manipulation

```{r}
mutate(litters_df, wt_gain = gd18_weight - gd0_weight)

mutate(litters_df, sq_pups = pups_born_alive^2)

mutate(litters_df, group = str_to_lower(group))

mutate(
  litters_df,
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group)
)
```

## `arrange`

Default ascending order but can change the arrangement to descending order using `desc` function. Arranging by multiple variables is also possible
```{r}
arrange(litters_df, gd0_weight)

arrange(litters_df, desc(gd0_weight))

arrange(litters_df, pups_born_alive, gd0_weight)
```

## `pipe`

The order should not matter as long as it doesn't depend on the previous variable
```{r}
litters_df = 
  read_csv("data/FAS_litters.csv", na = c("NA", "", ".")) %>% 
  janitor::clean_names() %>% 
  select(-pups_born_alive) %>% 
  filter(group == 'Con7') %>% 
  mutate(
    wt_gain = 
      gd18_weight - gd0_weight
  )

write_csv(litters_df, 'data/cleaned_FAS_litters.csv')
```


```{r}
read_csv("data/FAS_litters.csv", na = c("NA", ".", "")) %>% 
  janitor::clean_names() %>% 
  mutate(wt_gain = gd18_weight - gd0_weight) %>% 
  lm(wt_gain ~ pups_born_alive, data = )
```




